---
title: "Building Human Activity Recognition Model"
author: "Ivan Foong"
date: "19 June, 2014"
output: html_document
---

This document describe the process to explore and building a Human Activity Recognition Model based on the dataset available at <http://groupware.les.inf.puc-rio.br/har>, for my assignment on Practical Machine Learning course on Coursera.

Enable multicore processing to speed up training process

```{r,echo=FALSE}
# register multiple computational cores
library(doMC)
registerDoMC(cores = 7)
```

Download the csv data from the internet

```{r}
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv','pml-training.csv','curl')
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv','pml-testing.csv','curl')
```

Read in the data

```{r}
trainingDat <- read.csv('pml-training.csv')
testingDat <- read.csv('pml-testing.csv')
```

Exploring the data via scatter plots (Showing first four features as sample)

```{r, echo=FALSE}
require(gridExtra)
require(ggplot2)
plot1 <- ggplot(trainingDat, aes(x=classe, y=roll_belt, color=classe)) + geom_point(shape=1)
plot2 <- ggplot(trainingDat, aes(x=classe, y=pitch_belt, color=classe)) + geom_point(shape=1)
plot3 <- ggplot(trainingDat, aes(x=classe, y=yaw_belt, color=classe)) + geom_point(shape=1)
plot4 <- ggplot(trainingDat, aes(x=classe, y=total_accel_belt, color=classe)) + geom_point(shape=1)
grid.arrange(plot1,plot2,plot3,plot4)
```


Exploring the data further and found that the data has alot of NA

```{r, echo=FALSE}
#Number of valid values
is_not_na_sum <- sum(!is.na(trainingDat))

#Number of NA values
is_na_sum <- sum(is.na(trainingDat))

barplot(c(is_not_na_sum, is_na_sum), names.arg=c("Valid", "NA"), main="Valid vs NA values count for testing data")
```


Drop columns with NA values based on first testing set record, to speed up the training process

```{r}
usableVars <- !is.na(testingDat[1,])
trainingDat <- trainingDat[, usableVars]
testingDat <- testingDat[, usableVars]
```


Training a random forest model using training data, with 10 fold, 10 repeat, "out of bag" cross validation

```{r}
library(caret)
fitControl <- trainControl(method = "oob", number = 10, repeats = 10)
set.seed(1337)
fit <- train(classe ~ . - X - user_name - raw_timestamp_part_1 - raw_timestamp_part_2 - cvtd_timestamp - new_window - num_window, data = trainingDat, method = "rf", trControl = fitControl, verbose = FALSE)
fit
```


Check error rate of the trained model against training data, which has 0 error for this trained random forest model

```{r}
trainingDat.pred <- predict(fit,newdata=trainingDat)
trainingDat.error <- dim(trainingDat)[1] - sum(trainingDat$classe == trainingDat.pred)
trainingDat.error
```

Predict the testing data using the trained random forest model and assigning the prediction to testingDat$pred

```{r}
testingDat.pred <- predict(fit,newdata=testingDat)
testingDat.pred
```

